Embedded Scalars, Preferred Readings \& Prosody: {A}n Experimental Revisit

### 2015-02-04 text/paper/revision_01/paper.tex
K2 revision reflection finished

There are two main theoretical positions which make different
predictions about the readings of \as- and \es-sentences
\citep[see][for
overview]{Horn2006:The-Border-Wars,Geurts2010:Quantity-Implic,Sauerland2012:The-Computation,ChemlaSingh2014:Remarks-on-the-}. We
will refer to these as \mymark{traditionalism} and
\mymark{grammaticalism} and treat each one in turn.  As already
mentioned in the introduction, there is some leeway in assessing the
predictions for each approach depending on which auxiliary assumptions
to adopt. Therefore, we will distinguish different varieties of either
core theory. While there are certainly other varieties conceivable,
they are, in our perception, the most prominent ones, but, more
importantly, also the most practical ones in that they give us precise
and testable \emph{ex ante} predictions about expected behavioral data
(as opposed to \emph{ex post} explanations after inspecting the
data). The predictions of the considere core theories plus their
auxiliary assumptions are summarized in Table~\ref{tab:predictions} on
page~\pageref{tab:predictions}. Below, in the discussion section
(section~\ref{sec:discussion-1}), further conceivable varieties of
traditionalism and grammaticalism will be discussed \emph{after}
inspecting the relevant data.

@@

The foregoing interpretation and discussion presupposed that the
interpretation of our data, given in Section~\ref{sec:discussion-1},
was sound. But there are a number of conceivable objections to our
design and our interpretation of the presented results. We will
consider the most pressing ones in the following.

% To begin with, it could be questioned whether our design reliably
% indicates the preference relations in
% (\ref{bsp:observed-preferences}). One could object that it is not
% feasible to compare results from the preference-related controls to
% results from our target conditions, because the source of different
% readings is fundamentally different. In the former we have a syntactic
% ambiguity, in the latter a set of putative pragmatic enrichments. The
% former is a case of perceived ambiguity (some of our subjects reported
% this), the latter arguably is not
% \citep[see][]{GeurtsPouscoulous2009:Embedded-Implic}. So perhaps even
% if the distribution of responses is indicative of the known preference
% patterns in preference-related controls, this would not necessarily
% mean that the distribution of responses in target conditions is
% indicative of preference relations. Although we cannot rule out that
% the incremental verification task detects preferences among multiple
% readings only for some sources of variability in readings and not
% others, it is by no means clear why this should be so. The burden of
% argument is on the sceptic. But even if we assume that the incremental
% verification task is incapable of reliably indicating preferences
% among candidate pragmatic enrichments of the relevant kind, then it
% should at least be able to indicate availability of readings. So if we
% deny the observed response patterns the power to indicate preference,
% what should the information about availability of readings be that we
% may safely infer from it? It should be that local readings are
% available in \es-sentences and that global readings are unattested in
% \as- and \es-sentences. But that alone is not predicted by either
% traditionalism or grammaticalism. The former cannot explain the
% substantial number of local responses in \es-sentences, the latter
% does not explain the absence of total global responses.

% Another line of criticism that challenges the validity of
% (\ref{bsp:observed-preferences}) targets the Bayesian
% salience-competition model that we used to derive it. We readily admit
% that our model is highly simplistic. Several aspects are especially
% noteworthy here. Firstly, we did not distinguish different types of
% errors, and did not take into account that when errors occurred
% earlier in the sequence, later decision points could no longer be
% reached. This way the model may actually have unduly de-emphasized the
% salience of readings that can be judged later in a sequence. Another,
% more complicated model of the likelihood of answer patterns, might
% therefore have given support to different conclusions. Unfortunately,
% it is not clear how exactly such a more complex generating model
% should be set-up. Eventually, a full processing model, designed for
% data from incremental verification tasks, would be needed. We cannot
% offer such a model here, but can only the offer a first step in that
% direction. More importantly, though, the only part of
% (\ref{bsp:observed-preferences}) that might plausibly change for a
% more encompassing model along the lines sketched above is that the 3
% \emph{false} answers in the \as-conditions with neutral intonation
% might support the conclusion that the local reading is attested. But
% this hypothetical effect will certainly be no bigger than marginal,
% and so no genuinely different conclusions of theoretical relevance
% should be expected.

@@

We tested experimentally the predictions of traditionalism and
grammaticalism with respect to the availability of and preferences
over three types of conceivable pragmatic enrichments of two types of
sentences, in which scalar \emph{some} occurred in the scope of either
a monotonic quantifier \emph{all} or the non-monotonic quantifier
\emph{exactly one}. To avoid potential confounds of the pictorial
material, such as through typicality or contrast
\citep{Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Embedded-Scalar},
we employed an incremental verification task \citep{Conroy2008}, in
which subjects were presented with partially covered pictures and
asked to uncover the picture sequentially until they felt able to give
a categorical truth-value judgements. In order to control for
potential confounds of ``silent prosody,'' sentence material was
presented auditorily and contrastive stress on embedded \emph{some}
was manipulated. This way we were able to also explicitly test the
prosodic markedness hypothesis, which is often used to supplement
traditionalism. Additionally, we included preference-related control
conditions in order to be able to deduce latent preferences for
candidate readings from the answer patterns we gathered. To this end,
we introduced a generative Bayesian model that gives probabilistic
predictions about observable answer patterns in our task, given
concrete instances of relative saliences of readings and other
parameters. Using the observed data, in particular in connection with
the preference-related control conditions, we inferred a posteriori
likely values of the latent model parameters.

Backed up by our analyses, we concluded that both traditionalism and
grammaticalism do not predict our data well. Traditionalism fails to
account for high numbers of responses indicating local readings even
for prosodically unmarked \es-sentences. Grammaticalism, insofar as
it builds on disambiguation by logical strength, fails to predict the
absence of answers indicative of global readings and the relative
abundance of answers indicative of literal readings.

The clearest upshot of theoretical importance of this study is that
strength-based meaning selection does not seem to be a good
disambiguation criterion for grammaticalism. We tentatively suggested
an alternative selection principle for grammaticalism in terms of
``parsing economy.'' Another important conclusion from our data is
that \es-sentences, using non-monotonic quantifier \emph{exactly one},
appeared to be prone to receive a ``shallow interpretation.'' This may
be used by traditionalism to explain answers indicative of local
readings for \es-sentences. But the point is relevant in general. The
possibility of ``shallow interpretation'' of non-monotonic quantifiers
should be considered also when interpreting the results of other
experimental designs \citep[e.g.][]{CliftonDube2010:Embedded-Implic,
  ChemlaSpector2010:Experimental-Ev}.

We have argued that auditory presentation of sentence material should
be controlled for if prosodic information is hypothesized to influence
availability and preferences of readings. We also argued that
theoretical positions must be formulated in such a way that they make
testable predictions about reading preferences where different
potential readings are allowed (contra a grammaticalism without
commitment to any testable disambiguation principle). We should then
try to obtain information about actual preference patterns by explicit
comparison with structures whose preference structures are
well-studied, ideally with some model of the data generating process
for optimal data analysis.

>>>>

There are two main theoretical positions which make different
predictions about the readings of \as- and \es-sentences
\citep[see][for
overview]{Horn2006:The-Border-Wars,Geurts2010:Quantity-Implic,Sauerland2012:The-Computation,ChemlaSingh2014:Remarks-on-the-}. We
will refer to these as \mymark{traditionalism} and
\mymark{grammaticalism} and treat each one in turn.  As already
mentioned in the introduction, there is some leeway in assessing the
predictions for each approach depending on which auxiliary assumptions
to adopt. Therefore, we will distinguish different varieties of either
``core theory.'' While there are certainly other varieties
conceivable, they are, in our perception, the most prominent ones,
but, more importantly, also the most practical ones in that they give
us precise and testable \emph{ex ante} predictions about expected
behavioral data (as opposed to \emph{ex post} explanations after
inspecting the data). The predictions of the considered core theories
plus their auxiliary assumptions are summarized in
Table~\ref{tab:predictions} on page~\pageref{tab:predictions}. Below,
in Section~\ref{sec:critical-reflection}), further conceivable
varieties of traditionalism and grammaticalism will be discussed
\emph{after} inspecting the relevant data.

@@

The foregoing interpretation and discussion presupposed that the
interpretation of our data, given in Section~\ref{sec:discussion-1},
was sound. But there are a number of conceivable objections to our
design and our interpretation of the presented results, which
essentially all cast into doubt that our data suggests the preference
relation in (\ref{bsp:observed-preferences}). We will consider the
most pressing ones in the following.

\paragraph{The ``choose first match'' argument.} It is tempting to
think that participants generally tended to choose the first exit
possible. They might, after all, have simply wanted to save time and
be done with the experiment. We considered this a possible confound
and it is therefore that we included the preference-related
controls. But the objection is already clearly refuted by the data
itself. For it would entail that we should see many more global
readings for \es-sentences, which could be judged before the literal
readings, and also that we should have seen the reverse pattern in the
preference-related controls, where we did observe that the reading
that could be judged \emph{later} attracted comparatively more
responses.

\paragraph{Incomparability of target conditions and preference-related
  controls.} One could object that it is not feasible to compare
results from the preference-related controls to results from our
target conditions, because the source of different readings is
fundamentally different. In the former we have a syntactic ambiguity,
in the latter a set of putative pragmatic enrichments. The former is a
case of perceived ambiguity (some of our subjects reported this), the
latter arguably is not
\citep[see][]{GeurtsPouscoulous2009:Embedded-Implic}. So perhaps even
if the distribution of responses is indicative of the known preference
patterns in preference-related controls, this would not necessarily
mean that the distribution of responses in target conditions is
indicative of preference relations in the same way.

This objection is very serious. It is amplified by an even more
general worry, raised as a challenge by a reviewer, who intuits that
the incremental verification might altogether be insensitive to scalar
implicatures. According to the reviewer, it is to be expected that
since \emph{some} is typically strengthened in non-embedded
conditions, virtually all of our participants would exit at an early
position corresponding to the literal reading in a suitable version of
our \ivt.

To address this worry we ran another stripped-down version of the
\ivt. We recruited 50 subjects via Amazon's Mechanical Turk and paid
them 1 US\$ compensation. Similar to our main experiment, the
post-experiment started with two training trials with unambiguous
sentences, familiarizing our subjects with the task. In the
browser-based version, alert boxes pointed out subjects' mistakes when
they tried to exit to early, chose the wrong response or tried to
unravel the sequence further than strictly necessary. The main part of
the experiment consisted of six trials, two of which were critical and
four of which were controls. On critical trials, each subject saw
different version of sentences like (\ref{bsp:Item-PostExp}), in
connection with sequences like in Figure~\ref{fig:Trial-PostExp}.

\begin{exe}
  \ex \label{bsp:Item-PostExp} The scissors are connected to some of the circles.
\end{exe}

\begin{figure}[t]
	\centering
	\subfloat[][Step 2  (\lit true)]{ 
		\fbox{\includegraphics[width=3.5cm]{pictures/PostExp_2.pdf}}
	    \label{fig:exseqES1}
	}
	\subfloat[][Step 3]{
		\fbox{\includegraphics[width=3.5cm]{pictures/PostExp_3.pdf}}
	    \label{fig:exseqES2}
	}
	\subfloat[][Step 4 (\acro{prag} false)]{ 
		\fbox{\includegraphics[width=3.5cm]{pictures/PostExp_4.pdf}}
	    \label{fig:exseqES3}
	}
	\caption{Example sequence for critical trial in
          post-experiment.}
	\label{fig:Trial-PostExp}
\end{figure}

\noindent The first step in the sequence had all potential connections
covered, as before. On step 2, the critical sentence could be judged
true under a literal reading. On the last step 4, it could be judged
false under a pragmatically strengthened implicature reading. Control
conditions had unambiguous sentences and served to filter out subjects
with insufficient performance. From the total of 100 critical trials
(50 subjects times 2 trials each), 66 were literal answers, 27 were
pragmatic answers and 7 were `errors'. If we restrict attention to
only those participants that had at least two of the four control
conditions correct, we are left with 39 subjects. With these, we find
55 literal, 21 pragmatic and 2 `error' answers. In sum, this strongly
suggests that the \ivt is not generally insensitive to scalar
implicature and does give substantial counts also for implicature
readings if these can be judged later than a literal reading.

If we accept that the \ivt is not insensitive to implicature-based
ambiguities, what is left to worry about is that information about
response biases (early vs.~late) do not carry over from the
preference-related control conditions to the target conditions in the
main experiment. It could be, after all, that the task triggers
subjects to exit late for perceived syntactic ambiguities and to exit
early for ``pragmatic ambiguities'' involving implicature
readings. This might also possibly explain why there are more `error'
responses on the last position in the preference-related controls. But
none of this has any serious impact on the suggested preference
relation in (\ref{bsp:observed-preferences}). For even if, as possibly
the post-experiment suggests as well, there is a sequential bias in
the target conditions to exit early, the main effect of such a bias would
merely be that we should consider local readings more salient than we
do now. In the light of the data from main and post-experiments, it
would be far-fetched to uphold that our data indicates that local
readings are at least as salient as literal ones. But that means that
the suggested interpretation in (\ref{bsp:observed-preferences}) still
holds, and all the problems that we identified for (the considered
versions of) traditionalism and grammaticalism apply.

To strengthen this case further, let us assume that the \ivt is
entirely incapable of reliably indicating preferences among candidate
pragmatic enrichments of the relevant kind. Can we still draw any
conclusions about mere availability (non-negligible salience) of
readings? At a minimum, we would like to uphold that local readings
are available in \es-sentences and that global readings are unattested
in \as- and \es-sentences. But that alone is not predicted by either
traditionalism or grammaticalism in the guises that we considered. The
former cannot explain the substantial number of local responses in
\es-sentences, the latter does not explain the total absence of global
responses.

\paragraph{Interpretation of ``salience'' \& the Bayesian model.}
Another line of criticism that challenges the validity of
(\ref{bsp:observed-preferences}) targets the Bayesian
salience-competition model that we used to derive it. To avoid a
potential misunderstanding pointed out by a reviewer, it is not the
case the our interpretation of the \ivt assumes that each subject
fixes an interpretation at the beginning of the sequence and applies
it, unswayed by anything that is encountered along the sequence of
revealed pictures. The Bayesian model is compatible with this
interpretation, but it is also compatible with a another one that we
think is more plausible, namely that at each choice point during the
sequence of pictures, subjects' aggregate probabilistic behavior is
described as a function that compares the relative salience of the
readings available, together with potential response biases and
uniform error rates.

Still, we readily admit that our model is highly simplistic. Several
aspects are especially noteworthy here. Firstly, we did not
distinguish different types of errors, and did not take into account
that when errors occurred earlier in the sequence, later decision
points could no longer be reached. This way the model may actually
have unduly de-emphasized the salience of readings that can be judged
later in a sequence. Another, more complicated model of the likelihood
of answer patterns, might therefore have given support to
quantitatively different conclusions. Unfortunately, it is not clear
how exactly such a more complex generating model should be
set-up. Eventually, a full processing model, designed for data from
incremental verification tasks, would be needed. We cannot offer such
a model here, but can only the offer a first step in that
direction. More importantly, though, the only part of
(\ref{bsp:observed-preferences}) that might plausibly change for a
more encompassing model along the lines sketched above is that the 3
\emph{false} answers in the \as-conditions with neutral intonation
might support the conclusion that the local reading is attested. But
this hypothetical effect will certainly be no bigger than marginal,
and so no genuinely different conclusions of theoretical relevance
should be expected.

@@

We tested experimentally the predictions of suitably constrained
versions of traditionalism and grammaticalism with respect to the
availability of and preferences over three types of conceivable
pragmatic enrichments of two types of sentences, in which scalar
\emph{some} occurred in the scope of either a monotonic quantifier
\emph{all} or the non-monotonic quantifier \emph{exactly one}. To
avoid potential confounds of the pictorial material, such as through
typicality or contrast
\citep{Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Embedded-Scalar},
we employed an incremental verification task \citep{Conroy2008}, in
which subjects were presented with partially covered pictures and
asked to uncover the picture sequentially until they felt able to give
a categorical truth-value judgements. In order to control for
potential confounds of ``silent prosody,'' sentence material was
presented auditorily and contrastive stress on embedded \emph{some}
was manipulated. This way we were able to also explicitly test the
prosodic markedness hypothesis, which is often used to supplement
traditionalism. Additionally, we included preference-related control
conditions in order to be able to deduce latent preferences for
candidate readings from the answer patterns we gathered. To this end,
we introduced a generative Bayesian model that gives probabilistic
predictions about observable answer patterns in our task, given
concrete instances of relative saliences of readings and other
parameters. Using the observed data, in particular in connection with
the preference-related control conditions, we inferred a posteriori
likely values of the latent model parameters.

Backed up by our analyses, we concluded that both traditionalism and
grammaticalism do not predict our under the set of auxiliary
assumptions that we provided them with in
Section~\ref{sec:theories-predictions}. Traditionalism fails to
account for high numbers of responses indicating local readings even
for prosodically unmarked \es-sentences. Grammaticalism, insofar as it
builds on disambiguation by logical strength, fails to predict the
absence of answers indicative of global readings and the relative
abundance of answers indicative of literal readings.

The clearest upshot of theoretical importance of this study is that
strength-based meaning selection does not seem to be a good
disambiguation criterion for grammaticalism. This point is also
relevant in general, because disambiguation based on logical strength
has been suggested in other domains as well
\citep[e.g.][]{DalrympleKanazawa1998:Reciprocal-Expr,CobrerosEgre2012:Tolerant-Classi}.
We tentatively suggested an alternative selection principle for
grammaticalism in terms of ``parsing economy,'' but it remains to be
seen whether this is empirically successful in other domains as well.

Another important conclusion from our data is that \es-sentences,
using non-monotonic quantifier \emph{exactly one}, appeared to be
prone to receive an unexpected interpretation. This may be used by
traditionalism to explain answers indicative of local readings for
\es-sentences. But the point is relevant in general. The possibility
of unexpected interpretations of non-monotonic quantifiers should be
considered also when interpreting the results of other experimental
designs \citep[e.g.][]{CliftonDube2010:Embedded-Implic,
  ChemlaSpector2010:Experimental-Ev}.

We have argued that auditory presentation of sentence material should
be controlled for if prosodic information is hypothesized to influence
availability and preferences of readings. We also argued that
theoretical positions must be formulated in such a way that they make
testable predictions about reading preferences where different
potential readings are allowed. We should then try to obtain
information about actual preference patterns by explicit comparison
with structures whose preference structures are well-studied, ideally
with some model of the data generating process for optimal data
analysis.
